---
format: gfm
---

<!-- README.md is generated from README.qmd. Please edit that file -->

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.width = 6,
  out.width = "100%",
  fig.retina = 3
)
```

# ggsegThalamus

<!-- badges: start -->
[![R-CMD-check](https://github.com/ggsegverse/ggsegThalamus/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ggsegverse/ggsegThalamus/actions/workflows/R-CMD-check.yaml)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

Thalamic nuclei parcellations for the ggseg ecosystem.

Two published atlases of the thalamus, from different modalities:

| Atlas | Nuclei per hemisphere | Derived from |
|---|---|---|
| `thalamus_hcp()` | 7 | HCP diffusion MRI (Najdenovska et al. 2018) |
| `thalamus_thomas()` | 12 | White-matter-nulled MPRAGE (Su et al. 2019) |

THOMAS is the finer of the two, and the only one to separate the geniculate
nuclei, the habenula and the mammillothalamic tract.

Because these atlases subdivide the thalamus alone, the surrounding deep grey
structures are kept as grey anatomical context rather than cleared.

Labels keep the identifiers the published lookup tables use, `region` is those
stripped of the hemisphere and lower-cased, and a `name` column carries the
spelled-out nucleus name for printing.

For the FreeSurfer thalamic segmentation, a different atlas again, see
[`ggsegFreeSurfer`](https://github.com/ggsegverse/ggsegFreeSurfer).

## Atlas Citations

> Najdenovska E, Alemán-Gómez Y, Battistella G, et al. (2018). "In-vivo
> probabilistic atlas of human thalamic nuclei based on diffusion-weighted
> magnetic resonance imaging." *Scientific Data*, 5, 180270.
> DOI: [10.1038/sdata.2018.270](https://doi.org/10.1038/sdata.2018.270)

> Su JH, Thomas FT, Kasoff WS, et al. (2019). "Thalamus Optimized Multi Atlas
> Segmentation (THOMAS): fast, fully automated segmentation of thalamic nuclei
> from structural MRI." *NeuroImage*, 194, 272-282.
> DOI: [10.1016/j.neuroimage.2019.03.021](https://doi.org/10.1016/j.neuroimage.2019.03.021)

If you use this atlas in your work, please cite both the original atlas publication
and the ggseg package:

> Mowinckel AM, Vidal-Pineiro D (2020). "Visualization of Brain Statistics
> With R Packages ggseg and ggseg3d."
> *Advances in Methods and Practices in Psychological Science*, 3(4), 466-483.
> DOI: [10.1177/2515245920928009](https://doi.org/10.1177/2515245920928009)

## Installation

We recommend installing the ggseg-atlases through the ggsegverse
[r-universe](https://ggsegverse.r-universe.dev/#builds):

```{r}
#| label: install-r-universe
#| eval: false
options(repos = c(
  ggsegverse = "https://ggsegverse.r-universe.dev",
  CRAN = "https://cloud.r-project.org"
))

install.packages("ggsegThalamus")
```

You can install this package from [GitHub](https://github.com/) with:

```r
# install.packages("pak")
pak::pak("ggsegverse/ggsegThalamus")
```

## Usage

```{r}
#| label: plot-hcp
#| fig-height: 6
library(ggseg)
library(ggsegThalamus)

plot(thalamus_hcp())
```

The finer THOMAS segmentation:

```{r}
#| label: plot-thomas
#| fig-height: 6
plot(thalamus_thomas())
```

## Code of Conduct

Please note that the ggsegThalamus project is released with a
[Contributor Code of Conduct](https://contributor-covenant.org/version/2/1/CODE_OF_CONDUCT.html).
By contributing to this project, you agree to abide by its terms.
